### 2015-01-07 Studies/2012_PG_County/Report/PG_Co_Fire.tex
PG Co: rename year on PG County fire

To calculate the total fire size, the appropriate burning area of the basement needed to be determined. Based on the floor plan of the structure and post-incident images, two distinct fires were considered: 1 source in the basement bathroom and 1 source in the basement kitchen. The bathroom fire was estimated to be comprised of wood paneling along the four walls (17.3~m$^2$ or 186~ft$^2$) and wood subfloors and studs along the ceiling (2.9~m$^2$ or 31~ft$^2$). This results in 20.2~m$^2$ (217~ft$^2$) of wood burning surface area at 50~kW/m$^2$ or a 1~MW energy release rate. The remainder of the wood combustable load in the basement was lumped into the localized kitchen fire. The ceiling and wall areas were included in the basement kitchen and main living area totally approximately 86~m$^2$ (926~ft$^2$). This would result in a 4.3~MW fire.

Initial simulations were conducted with the two point location fires totaling 5.3~MW, however, the fire behavior did not align with observations made during the fire incident or post-incident images. Specifically there was insufficient fuel in the basement to drive the basement to an underventilated condition. At the time of the fire the structure was unoccupied~\cite{PGCounty2013}, there were additional fuels that were not accounted for: transient fuels, sparse kitchen cabinetry, and an unknown bathroom vanity among other fuels. A second reason is that the 50~kW/m$^2$ HRRPUA used to estimate the first size does not account for an observed increase in burning rate during wind driven fires~\cite{madrzykowski2009fire,kerber2009fire}. It is not clear how to quantify the impact of wind on burning rate within a structure from first principles, so several iterations of the model with increased fire sizes were run. The final estimated fire size was a 1.8~MW fire in the bathroom and a 6.75~MW fire in the kitchen. This resulted in a total peak specified fire size of 8.55~MW.

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The fires started at the beginning of the simulation and increased to a value of 8.5~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. Since the actual ignition time is unknown, the fire was ramped quickly to reflect the conditions determined from reports and post-incident images. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations in the basement (1 source for the kitchen, 1 source for the bathroom).  At 207~s, using a simple decay model, the fire size decreases linearly to 25~\% of its peak value in 60~s. This is to account for the impact of firefighters putting water on the basement fire (Table~\ref{tab:fire_info}). The following lines define the source fires in the FDS input file:

@@

Figure~\ref{fig:hrr} shows the overall prescribed design fire for this scenario, which is the sum of the HRRs of the three source fires. In this figure, the text labels the prescribed steady peak HRR and the decay phase associated with firefighter water application. Therefore, the estimated maximum specified fire size for this structure was approximately 8.5~MW. Note that the prescribed HRR can be different than the simulated HRR calculated by FDS based on the amount of available fuel, oxygen, and ventilation. A comparison of the prescribed HRR and the simulated HRR is discussed in Section~\ref{HRR}.

@@

\section{Materials}
\label{matl}
While the fires were represented as fuel sources with a constant area (Section~\ref{fire}), from a heat transfer perspective, it is important to define the material properties (density, thermal conductivity, and specific heat) to account for heat transfer and energy storage in the ceiling, walls, and floor. In this study, the material properties of gypsum board~\cite{WAKILI2007} were specified on the finished walls and ceilings on the main two floors of the structure, and the material properties of wood~\cite{Incropera:1} were specified on the unfinished portions of the attic and rear porch. The following lines define these materials in the FDS input file:

@@

\section{Ventilation}
\label{Vents}
The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors, breaking windows), fire acting on the structure (fire breaching windows/doors) and wind (restricting flow, closing doors). The ventilation times and ventilation areas represent our best understanding of the incident. The times of ventilation changes for doors and windows are provided in Table~\ref{tab:vents}.

@@

Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open rear basement door and two open windows at the start of the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

@@

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condition, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompasses the structure with a total volume 14~m (46~ft) by 12~m (39ft) by 8~m (26~ft) that required a total of 1.3 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. The 20~cm grid resolution extends the computational domain 5~m to the left and right sides and 4~m behind the 10~cm mesh to provide sufficient computational domain for the wind profile to develop around the structure. The mesh is coarser to cut down on computational time and because high accuracy in the fluid mechanics of the air flow far from the structure is not important in the analysis of this simulation. The 20~cm resolution resulted in an additional 216,000 computational cells.

Figure~\ref{fig:geom_grid} shows the front and rear of the structure and surrounding domain rendered in Smokeview with the 10~cm and 20~cm meshes. Note the finer mesh (smaller cells) surrounding the structure and the coarser mesh on each side. The domain was divided into 12 equally sized 10~cm meshes that surround the structure (each containing 150,000 grid cells) plus 4 equally sized 20~cm meshes to the left and right of the structure (each containing 32,000 grid cells) and one 20~cm mesh in the rear of the structure (containing 56,000 grid cells) for a total of 17 meshes. Multiple meshes allowed the simulation to be processed in parallel, which reduced the amount of required calculation time to approximately 3~days for the 20~mph wind case. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. The boundary between the twelve 10~cm meshes and five 20~cm meshes is indicated by the {\bf bold} lines. Table~\ref{tab:mod_param} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

Figure~\ref{fig:PG_Total_HRR} shows the HRR in the structure vs. time based on the fires described in Section~\ref{fire}. In this figure, the solid line represents the prescribed HRR that was input into the simulation (based on a prescribed mass flux), and the dashed line represents the HRR inside the structure that was calculated by FDS based on the ventilation conditions. The vertical dashed lines represent the times at which ventilation occurred: front door opening and closing as well as the breaking of a front window for emergency escape. See Table~\ref{tab:vents} for a detailed list of the ventilation that occurring during the incident. The maximum size of the fire specified in this study was selected such that the conditions aligned with observations and fuel loads from the actual fire incident.

@@

In Fig.~\ref{fig:PG_Total_HRR}, the prescribed HRR that was input into the simulation is different than the HRR calculated by FDS based on the ventilation conditions. From 0~s to approximately 9~s, the calculated HRR was in agreement with the prescribed HRR because there was an adequate amount of oxygen for all of the fuel to combust. The differences in prescribed versus predicted HRR is a result of either fuel within the structure accumulating and not combusting because of a lack of oxygen or combustion occurring exterior to the structure where fuel and oxygen can mix and react. Fig.~\ref{fig:PG_9MW_HRR} shows the total calculated HRR within the computational domain (includes combustion occurring exterior to the structure) compared to the prescribed HRR curve.

@@

A comparison of the differences between the HRR inside the structure (Fig.~\ref{fig:PG_Total_HRR}) and the total HRR (Fig.~\ref{fig:PG_9MW_HRR}) over the first 100 seconds of the simulation indicates that on average there was the equivalent of an approximate 0.675 MW fire(s) occurring through open external vents. This aligns with observed post-fire damage and post-incident report~\cite{PGCounty2013}. Fig.~\ref{fig:ext_fire_dam} shows a post-fire image of the back left corner of the structure as well as a Smokeview rendering of the fire from the FDS simulation prior to the front door being opening by firefighters.

@@

The growth of the fire combined with a 8.9~m/s (20~mph) wind flowing normal to the rear of the structure caused the pressure to rise within the structure. Hot gases which may have built up within the structure will flow from high pressure to low pressure. Based on the event timeline, the front door was opened 100 seconds into the simulation, which would create a low pressure relief on the first floor. Due to firefighters entering the structure through the front door, the simulated pressure in the first floor prior to and just after opening the door is important. Figure~\ref{fig:pressure_slices} shows the calculated pressure conditions in the structure just before and after the front door was opened (100~s) in the simulation.

In the top snapshot in Fig.~\ref{fig:pressure_slices}, the pressure profile shows an overpressure of approximately 200~Pa (2.9 $\times$ 10$^{-2}$~psi) prior to the door being opened. This pressure may be on the conservative side (over-prediction) because of wind speed fluctuations, wind direction, and small amounts of structure leakage. One second after the door was opened (bottom snapshot in Fig.~\ref{fig:pressure_slices}), the pressure on the first floor drops to between 75~Pa (1.1 $\times$ 10$^{-2}$~psi) and 85~Pa (1.2 $\times$ 10$^{-2}$~psi). Without the presence of wind, the pressure in the first floor prior to the door opening is is approximately 18~Pa (2.6 $\times$ 10$^{-3}$~psi) and after the door is opened the pressure drops to between 5~Pa (7.3 $\times$ 10$^{-4}$~psi) and 10~Pa (1.4 $\times$ 10$^{-3}$~psi).

@@

Analysis of the temperatures from the simulation focuses on the high-hazard area of the structure: the first floor area between the front door and the top of the interior stairwell. The temperature contours shown in Fig.~\ref{fig:temperature_slice_99s} indicate the gas temperatures within the structure 1~s before the front door was opened 0.91~m (3~ft) above the floor. The overlaid arrows represent the gas flow: the length of the lines represents speeds and the direction of the arrow represents the flow direction. Prior to the door opening, the open living area and kitchen on the first floor were approximately 60~$^{\circ}$C (140~$^{\circ}$F) while the peak temperatures were concentrated near the interior stairwell, 105~$^{\circ}$C (220~$^{\circ}$F). Note that there is very minimal flow throughout the first floor. Figures~\ref{fig:temperature_slice_101s} and \ref{fig:temperature_slice_105s} show gas temperature contours at 0.91~m (3~ft) above the floor, 1~s and 5~s after the front door was opened.

@@

After the front door is opened, hot fire gases from the basement begin to flow through the interior stairwell into the first floor. Figure~\ref{fig:temperature_slice_101s}, 1~s after the front door was opened, shows that the temperature contours at the top of the stairwell have increased to approximately 200~$^{\circ}$C (392~$^{\circ}$F) with a hot spot that approaches 425~$^{\circ}$C (800~$^{\circ}$F). Figure~\ref{fig:temperature_slice_105s}, 5~s after the front door was opened, shows more significant increases in temperature. The area of at the top of the stairs is most between 375~$^{\circ}$C (700~$^{\circ}$F) and 425~$^{\circ}$C (800~$^{\circ}$F). The gas temperatures 0.91~m (3~ft) above the floor at the front door have also increased to approximately 230~$^{\circ}$C (450~$^{\circ}$F) and show increased flow velocities.

@@

The report from Prince George's County Fire/EMS~\cite{PGCounty2013} indicated that two firefighters from Truck 809 were injured after making entry to the first floor and had to either bail out through a window or be rescued through the front door. From Section~\ref{temp}, the conditions on the first floor were initially tenable. After the front door was opened, however, the simulation results indicate that a high hazard area existed in the first floor that exceeded the conditions of a Class~III exposure (temperatures greater than 260~$^{\circ}$C or 500~$^{\circ}$F)~\cite{Donnelly2006}.

The simulation results indicate that a flow path was established between the open vents (windows and door) on the rear side of the basement and front side of the structure (through the front door of the first floor) after the front door was opened. Strong winds (8.9~m/s (20~mph)) pushed air into the rear side of the basement through window failures and combined with buoyant combustion gases to pressurize the structure. Opening the front door provided a pressure relief; this resulted in a rapid change in the conditions within the structure and the establishment of a flow path. After the front door was opened, the combustion gases at elevated temperature and pressure in the basement flowed towards lower pressure regions outside of the structure on the leeward side (front of structure) via the interior stairwell. The two injured firefighters were located in the flow path on first floor between the top of the interior stairwell and the front door.

To examine the impact of opening the front door on the establishment of a flow path, flow conditions are examined along the centerline of the interior stairwell doorway and at the front door. Figure~\ref{fig:flow_path_99s} shows the flow vectors at the first floor stairwell doorway 1 second before the front door was opened. Note that length and direction of the arrow indicates the magnitude and direction of the flow respectively. The color of the arrow corresponds to the values from the included temperature scales. Therefore, a short blue arrow represents slow moving, cool gas while a long red arrow represents fast moving, high temperature gas. Prior to the door opening, there is minimal flow through the first floor stairwell door and the temperature peak is approximately 125~$^{\circ}$C (257~~$^{\circ}$F) as shown in Figure~\ref{fig:flow_path_99s}. One second after the door opened, Figure~\ref{fig:flow_path_101s}, there is noticeable flow through both the stairwell doorway and through the front door. Simulation temperatures at the stairwell increased to approximately 200~$^{\circ}$C (392~~$^{\circ}$F) and temperatures through the doorway are approximately 100~$^{\circ}$C (212~~$^{\circ}$F). These values are also shown in planar contours 1~m above the floor in Figs.~\ref{fig:velocity_slices} and \ref{fig:temperature_slice_101s} for velocity and temperature, respectively.

@@

At the estimated time of firefighter entry to the first floor (142~s into the simulation; 21:14:42 according to Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. Velocities through both the doorway at the top of the stairs and the front door were approximately 13~m/s (29~mph), faster than the incident wind  speed (9~m/s or 20~mph) because of the addition of the buoyancy of the fire plume. The model temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750-800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against the front door 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.

@@

Figures~\ref{fig:int_temp_99s} to \ref{fig:int_temp_142s} show two interior views of temperature contours at particular snapshots in time. The top image is a planar view through the middle of the front door (from the front to the back of the house) which shows the conditions in the living area on the first floor and the conditions in the basement just past the bottom of the stairs. The bottom image shows temperature contours through the middle of the interior stairwell which connected the basement to the first floor of the structure (looking from the rear of the structure towards the front). Prior to the front door being opened (Fig.~\ref{fig:int_temp_99s}, temperatures in excess of 200~$^{\circ}$C (400~$^{\circ}$F) were confined to basement of the structure: specifically the rear of the structure and under the stairwell. One second after the door was opened, hazard in the structure increased within the structure with the most drastic change occurring at the top of the stairwell as peak simulated temperatures exceeded 400~$^{\circ}$C (750~$^{\circ}$F). At the approximate time firefighters made entry to the first floor, Fig.~\ref{fig:int_temp_142s}, the conditions within the flowpath (interior stairwell through front door) had become more severe with temperatures in excess of 400~$^{\circ}$C (750~$^{\circ}$F), which exceeds the Class III exposure of 260~$^{\circ}$C (500~$^{\circ}$F). These high-hazard conditions within the flowpath remained until the model HRR was reduced to simulate the action of applying water the fire.

@@

\chapter{Summary}
Fire Dynamics Simulator was used to provide insight into the fire dynamics of a fire that occurred within a single-story plus basement, single-family residential structure in Riverdale Heights, MD, that resulted in the serious injury of two firefighters. The fuel, fire size, and fire growth rate that were used in the FDS simulation were estimated by taking into account all of the available information including the post incident report from Prince George's Country Fire/EMS~\cite{PGCounty2013}, post-incident pictures, and relevant literature. This resulted in a maximum specified source fire of approximately 8.5~MW in the basement. Based on the limited ventilation conditions in the basement, the FDS simulation results indicated that before the front door was opened, there was combustion occurring outside the structure at the rear basement windows which had failed prior to arrival. Once the front door was opened, there was sufficient oxygen for combustion to occur within the structure, which intermittent flaming combustion on the first floor.

>>>>

To calculate the total fire size, the appropriate burning area of the basement needed to be determined. Based on the floor plan of the structure and post-incident images, two distinct fires were considered: 1 source in the basement bathroom and 1 source in the basement kitchen. The bathroom fire was estimated to be comprised of wood paneling along the four walls (17.3~m$^2$ or 186~ft$^2$) and wood subfloors and studs along the ceiling (2.9~m$^2$ or 31~ft$^2$). This resulted in 20.2~m$^2$ (217~ft$^2$) of wood burning surface area or a 1~MW fire using a 50~kW/m$^2$ HRRPUA. The remainder of the wood combustable load in the basement was lumped into the localized kitchen fire. The ceiling and wall areas included in the basement kitchen and main living area totaled approximately 86~m$^2$ (926~ft$^2$). This resulted in a 4.3~MW fire.

Initial simulations were conducted using the two fires totaling 5.3~MW, however, the fire behavior did not align with observations made during the fire incident or post-incident images. Specifically there was insufficient fuel to drive the basement to an underventilated condition. One reason is that while the structure was unoccupied at the time of the fire~\cite{PGCounty2013}, there were additional fuels that were not accounted for: transient fuels, sparse kitchen cabinetry, and an unknown bathroom vanity among other fuels. A second reason is that the 50~kW/m$^2$ HRRPUA used to estimate the fire size does not account for an observed increase in burning rate during wind driven fires~\cite{madrzykowski2009fire,kerber2009fire}. It is not clear how to quantify the impact of wind on burning rate within a structure from first principles, so several iterations of the model with increased fire sizes were run. The final estimated fire size was a 1.8~MW fire in the bathroom and a 6.75~MW fire in the kitchen. This resulted in a total peak specified fire size of 8.55~MW.

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The fires started at the beginning of the simulation and increased to a value of 8.55~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. Since the actual ignition time is unknown, the fire was ramped quickly to reflect the conditions determined from reports and post-incident images. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations in the basement (1 source for the kitchen, 1 source for the bathroom).  At 207~s, using a simple decay model, the fire size decreases linearly to 25~\% of its peak value in 60~s. This is to account for the impact of firefighters putting water on the basement fire (Table~\ref{tab:fire_info}). The following lines define the source fires in the FDS input file:

@@

Figure~\ref{fig:hrr} shows the overall prescribed design fire for this scenario, which is the sum of the HRRs of the two source fires. In this figure, the text indicates the prescribed steady peak HRR and the decay phase associated with firefighter water application. Note that the prescribed HRR can be different than the simulated HRR calculated by FDS based on the amount of available fuel, oxygen, and ventilation. A comparison of the prescribed HRR and the simulated HRR is discussed in Section~\ref{HRR}.

@@

\section{Materials}
\label{matl}
While the fires were represented as fuel sources with a constant area (Section~\ref{fire}), from a heat transfer perspective, it is important to define the material properties (density, thermal conductivity, and specific heat) to account for heat transfer and energy storage in the ceiling, walls, and floor. In this study, the material properties of gypsum board~\cite{WAKILI2007} were specified on the finished walls and ceilings on the main two floors of the structure, and the material properties of wood~\cite{Incropera:1} were specified on the unfinished portions of the structure. The following lines define these materials in the FDS input file:

@@

\section{Ventilation}
\label{Vents}
The simulation in this study accounts for changes in the ventilation due to a combination of fire department operations (opening doors), fire acting on the structure (fire breaching windows) and wind (restricting flow, closing doors). The ventilation times and ventilation areas represent the best understanding of the incident. The times of ventilation changes for doors and windows are provided in Table~\ref{tab:vents}.

@@

Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open rear basement windows at the start of the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

@@

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condition, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompassed the structure with a total volume 14~m (46~ft) by 12~m (39ft) by 8~m (26~ft) that required a total of 1.3 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. The 20~cm grid resolution extended the computational domain 5~m to the left and right sides and 4~m behind the 10~cm mesh to provide sufficient computational domain for the wind profile to develop around the structure. The mesh is coarser to cut down on computational time and because high accuracy in the fluid mechanics of the air flow far from the structure is not important in the analysis of this simulation. The 20~cm resolution resulted in an additional 216,000 computational cells.

Figure~\ref{fig:geom_grid} shows the front and rear of the structure and surrounding domain rendered in Smokeview with the 10~cm and 20~cm meshes. Note the finer mesh (smaller cells) surrounding the structure and the coarser mesh on each side. The domain was divided into 12 equally sized 10~cm meshes that surrounded the structure (each containing 150,000 grid cells) plus 4 equally sized 20~cm meshes to the left and right of the structure (each containing 32,000 grid cells) and one 20~cm mesh in the rear of the structure (containing 56,000 grid cells) for a total of 17 meshes. Multiple meshes allowed the simulation to be processed in parallel, which reduced the amount of required calculation time to approximately 3~days from over 1~week with a uniform 10~cm mesh. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. The boundary between the twelve 10~cm meshes and five 20~cm meshes is indicated by the {\bf bold} lines.

@@

Figure~\ref{fig:PG_Total_HRR} shows the HRR in the structure vs. time based on the fires described in Section~\ref{fire}. In this figure, the solid line represents the prescribed HRR that was input into the simulation (based on a prescribed mass flux), and the dashed line represents the HRR inside the structure that was calculated by FDS based on the ventilation conditions. The vertical lines represent the times at which ventilation occurred: front door opening and closing as well as the breaking of a front window for emergency escape. See Table~\ref{tab:vents} for a detailed list of the ventilation that occurring during the incident. The maximum size of the fire specified in this study was selected such that the conditions aligned with observations and fuel loads from the actual fire incident.

@@

In Fig.~\ref{fig:PG_Total_HRR}, the prescribed HRR that was input into the simulation is different than the HRR calculated by FDS based on the ventilation conditions. From 0~s to approximately 9~s, the calculated HRR was in agreement with the prescribed HRR because there was an adequate amount of oxygen for all of the fuel to combust. The differences in prescribed versus predicted HRR is a result of either fuel accumulation within the structure not combusting because of a lack of oxygen or combustion occurring exterior to the structure where fuel and oxygen can mix and react. Fig.~\ref{fig:PG_9MW_HRR} shows the total calculated HRR within the computational domain (includes combustion occurring exterior to the structure) compared to the prescribed HRR curve.

@@

A comparison of the differences between the HRR inside the structure (Fig.~\ref{fig:PG_Total_HRR}) and the total HRR (Fig.~\ref{fig:PG_9MW_HRR}) over the first 100 seconds of the simulation indicated that on average there was the equivalent of an approximate 0.675 MW fire(s) occurring through open external vents (basement windows). This aligns with observed post-fire damage and post-incident report~\cite{PGCounty2013}. Fig.~\ref{fig:ext_fire_dam} shows a post-fire image of the back left corner of the structure as well as a Smokeview rendering of the fire from the FDS simulation prior to the front door being opening by firefighters.

@@

The growth of the fire combined with a 8.9~m/s (20~mph) wind flowing normal to the rear of the structure caused the pressure to rise within the structure. Hot gases which may have built up within the structure will flow from high pressure to low pressure. Based on the event timeline, the front door was opened 100 seconds into the simulation, which would create a low pressure relief on the first floor. Due to firefighters entering the structure through the front door, the simulated pressure on the first floor prior to and just after opening the door is important. Figure~\ref{fig:pressure_slices} shows the calculated pressure conditions in the structure just before and after the front door was opened (100~s) in the simulation.

In the top snapshot in Fig.~\ref{fig:pressure_slices}, the pressure profile shows an overpressure of approximately 200~Pa (2.9 $\times$ 10$^{-2}$~psi), 1~s prior to the door being opened. This pressure may be on the conservative side (over-prediction) because of wind speed fluctuations, wind direction, and small amounts of structure leakage. One second after the door was opened (bottom snapshot in Fig.~\ref{fig:pressure_slices}), the pressure on the first floor drops to between 75~Pa (1.1 $\times$ 10$^{-2}$~psi) and 85~Pa (1.2 $\times$ 10$^{-2}$~psi). Without the presence of wind, the pressure in the first floor prior to the door opening is is approximately 18~Pa (2.6 $\times$ 10$^{-3}$~psi) and after the door is opened the pressure drops to between 5~Pa (7.3 $\times$ 10$^{-4}$~psi) and 10~Pa (1.4 $\times$ 10$^{-3}$~psi).

@@

Analysis of the temperatures from the simulation focuses on the high-hazard area of the structure: the first floor area between the front door and the top of the interior stairwell. The temperature contours shown in Fig.~\ref{fig:temperature_slice_99s} indicate the gas temperatures within the structure 1~s before the front door was opened 0.91~m (3~ft) above the floor. The overlaid arrows represent the gas flow: the length of the lines represents the magnitude of the speed and the direction of the arrow represents the flow direction. Prior to the door opening, the open living area and kitchen on the first floor were approximately 60~$^{\circ}$C (140~$^{\circ}$F) while the peak temperatures were concentrated near the interior stairwell, 105~$^{\circ}$C (220~$^{\circ}$F). Note that there is very minimal flow throughout the first floor. Figures~\ref{fig:temperature_slice_101s} and \ref{fig:temperature_slice_105s} show gas temperature contours at 0.91~m (3~ft) above the floor, 1~s and 5~s after the front door was opened.

@@

After the front door is opened, hot fire gases from the basement begin to flow through the interior stairwell into the first floor. Figure~\ref{fig:temperature_slice_101s}, 1~s after the front door was opened, shows that the temperature contours at the top of the stairwell have increased to approximately 200~$^{\circ}$C (392~$^{\circ}$F) with a hot spot that approached 425~$^{\circ}$C (800~$^{\circ}$F). Figure~\ref{fig:temperature_slice_105s}, 5~s after the front door was opened, shows more significant increases in temperature. The area of at the top of the stairs is between 375~$^{\circ}$C (700~$^{\circ}$F) and 425~$^{\circ}$C (800~$^{\circ}$F). The gas temperatures 0.91~m (3~ft) above the floor at the front door have also increased to approximately 230~$^{\circ}$C (450~$^{\circ}$F) and show increased flow velocities.

@@

The report from Prince George's County Fire/EMS~\cite{PGCounty2013} indicated that two firefighters from Truck 809 were injured after making entry to the first floor and had to either bail out through a window or be rescued through the front door. From Section~\ref{temp}, the conditions on the first floor were initially tenable. After the front door was opened, however, the simulation results indicated that a high hazard area existed in the first floor that exceeded the conditions of a Class~III exposure (temperatures greater than 260~$^{\circ}$C or 500~$^{\circ}$F)~\cite{Donnelly2006}.

The simulation results indicate that a flow path was established between the open vents (windows) on the rear side of the basement and front side of the structure (through the front door of the first floor) after the front door was opened. Strong winds (8.9~m/s (20~mph)) pushed air into the rear side of the basement through window failures and combined with buoyant combustion gases to pressurize the structure. Opening the front door provided a pressure relief; this resulted in a rapid change in the conditions within the structure and the establishment of a flow path. After the front door was opened, the combustion gases at elevated temperature and pressure in the basement flowed towards lower pressure regions outside of the structure on the leeward side (front of structure) via the interior stairwell. The two injured firefighters were located in the flow path on first floor between the top of the interior stairwell and the front door.

To examine the impact of opening the front door on the establishment of a flow path, flow conditions are examined along the centerline of the interior stairwell doorway and through the centerline of front door. Figure~\ref{fig:flow_path_99s} shows the flow vectors at the first floor stairwell doorway 1~s before the front door was opened. Note that length and direction of the arrow indicates the magnitude and direction of the flow respectively. The color of the arrow corresponds to the values from the included temperature scales. Therefore, a short blue arrow represents slow moving, cool gas while a long red arrow represents fast moving, high temperature gas. Prior to the door opening, there was minimal flow through the first floor stairwell door and the temperature peak is approximately 125~$^{\circ}$C (257~~$^{\circ}$F) as shown in Figure~\ref{fig:flow_path_99s}. One second after the door opened, Figure~\ref{fig:flow_path_101s}, there is noticeable flow through both the stairwell doorway and through the front door. Simulation temperatures at the stairwell increased to approximately 200~$^{\circ}$C (392~~$^{\circ}$F) and temperatures through the doorway are approximately 100~$^{\circ}$C (212~~$^{\circ}$F). These values are also shown in planar contours 1~m above the floor in Figs.~\ref{fig:velocity_slices} and \ref{fig:temperature_slice_101s} for velocity and temperature, respectively.

@@

At the estimated time of firefighter entry to the first floor (142~s into the simulation; 21:14:42 according to Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. Velocities through both the doorway at the top of the stairs and the front door were approximately 13~m/s (29~mph), faster than the imposed wind speed (9~m/s or 20~mph) because of the addition of the buoyancy of the fire plume. The model temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750-800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against the front door 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, the temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.

@@

Figures~\ref{fig:int_temp_99s} to \ref{fig:int_temp_142s} show two interior views of temperature contours at particular snapshots in time. The top image is a planar view through the middle of the front door (from the front to the back of the house) which shows the conditions in the living area on the first floor and the conditions in the basement just past the bottom of the stairs. The bottom image shows temperature contours through the middle of the interior stairwell which connected the basement to the first floor of the structure (looking from the rear of the structure towards the front). Prior to the front door being opened, (Fig.~\ref{fig:int_temp_99s} shows that temperatures in excess of 200~$^{\circ}$C (400~$^{\circ}$F) were confined to basement of the structure: specifically the rear of the structure and under the stairwell. One second after the door was opened, hazard in the structure increased within the structure, the most drastic change occurring at the top of the stairwell as peak simulated temperatures exceeded 400~$^{\circ}$C (750~$^{\circ}$F). At the approximate time firefighters made entry to the first floor, Fig.~\ref{fig:int_temp_142s}, the conditions within the flowpath (interior stairwell through front door) had become more severe with temperatures in excess of 400~$^{\circ}$C (750~$^{\circ}$F), which exceeds the Class III exposure of 260~$^{\circ}$C (500~$^{\circ}$F). These high-hazard conditions within the flowpath remained until the modeled HRR was reduced to simulate the action of applying water the fire.

@@

\chapter{Summary}
Fire Dynamics Simulator was used to provide insight into the fire dynamics of a fire that occurred within a single-story, single-family residential structure in Riverdale Heights, MD, that resulted in the serious injury of two firefighters. The fuel, fire size, and fire growth rate that were used in the FDS simulation were estimated by taking into account all of the available information including the post incident report from Prince George's Country Fire/EMS~\cite{PGCounty2013}, post-incident pictures, and relevant literature. This resulted in a maximum specified source fire of approximately 8.5~MW in the basement. Based on the limited ventilation conditions in the basement, the FDS simulation results indicated that before the front door was opened, there was combustion occurring outside the structure at the rear basement windows which had failed prior to fire department arrival. Once the front door was opened, there was sufficient oxygen for combustion to occur within the structure, with intermittent flaming combustion on the first floor.
